### 2010-07-09 doc/guide/aggregates.tex
git-svn-id: https://potassco.svn.sourceforge.net/svnroot/potassco/trunk/gringo@3171 3e02f596-3131-0410-8352-e04eeaec37cc

\com{Rewrote much here}

An aggregate is an operation on a multiset of weighted literals that evaluates to a value,
depending on the type of aggregate this might be integers but, 
e.g., \const{\#min} and \const{\#max} aggregates also permit function symbols.
In combination with comparisons, we can extract a truth value from
an aggregate's evaluation, thus, obtaining an aggregate atom.
Customizing the notation in~\cite{ferraris05a},
we consider ground \emph{aggregate atoms} of the form:
%
\begin{equation}\label{eq:aggregate:atom}
  l~ 
  \mathrm{op} 
  \code{[}
  L_1\code{=}w_1\code{,}
  \dots\code{,}
  L_n\code{=}w_n
  \code{]}
  u
\end{equation}
%
where~$l$ and~$u$ are values each of which can possibly be omitted,
$\mathrm{op}$ is a function from multisets of values to a value,
each $L_i$ is a literal of the form~$A$ or $\code{not}~A$
for some atom~$A\in\ground{\mathcal{A}}$,
and each~$w_i$ is a value.
The intuitive reading of an aggregate atom is that the value returned by~$\mathrm{op}$
applied to all weights~$w_i$ such that~$L_i$ holds should be in-between
lower bound~$l$ and upper bound~$u$ (inclusively).
If~$l$ or~$u$ is omitted,
it means that the aggregate's value is not constrained from below or above,
respectively.

For an aggregate atom~$\mathit{Agg}$ as in Equation(\ref{eq:aggregate:atom}),
we let $\atom{\mathit{Agg}}=\{L_i\in\nolinebreak\ground{\mathcal{A}}\mid 1\leq i\leq n\}$
denote the set of atoms occurring positively in~$\mathit{Agg}$.
Currently, \gringo\ and \clingo\ support aggregate operations
\const{\#sum}, \const{\#min}, and \const{\#max}, \const{\#avg}, \const{\#even}, and \const{\#odd}.%
\footnote{%
  The additional operation \const{count} is a shorthand for a \const{sum}
  aggregate in which all weights are~\const{1}.}
We can now extend translations~$\phi$ and~$\psi$ from Section~\ref{subsec:semantics}
to aggregate atoms, 
using the approach in~\cite{ferraris05a} but
applying the ``choice semantics'' for aggregates
in heads of rules~\cite{ferlif05a,siniso02a}:%

\newcommand\shiftleft{\kern\leftshift&\kern-\leftshift}

\newlength{\leftshift} 
\setlength{\leftshift} {2em}
%\settowidth{\leftshift}{$\phi^{+}[ l~ \const{\#sum} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]} ]$}

\begin{align}
  % l op u (head) %
  \shiftleft
  \psi[ \lhd~ \mathrm{op} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]}\rhd ] 
  \label{eq:aggr_head} \\
   &= \phi[ \triangleleft~ \mathrm{op} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]}\triangleright ] \notag \\
   &\wedge \bigl(\bigwedge\nolimits_{A\in atom(\triangleleft~ \mathrm{op} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]}\triangleright)}(A\vee\neg A)\bigr) \notag \\
  % l op u %
  \shiftleft
  \phi[ l~ \mathrm{op} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]} u ]
  \label{eq:aggr_lu} \\
  &= \phi[ l~ \mathrm{op} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]} ] \wedge{} 
    \phi[ \mathrm{op} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]} u ] \notag \\
  % op %
  \shiftleft
  \phi[ \mathrm{op} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]} ] = \top
  \label{eq:aggr}
\end{align}
Translation~$\psi$ of an aggregate atom in the head of a rule consists of two parts:
first, $\phi$ is used as defined in the following
(in Equation (\ref{eq:aggr_head}), $\lhd$ and~$\rhd$ are used as ``wildcards'' for provided or omitted 
 lower and upper bounds),
second, a disjunction $(A\vee\neg A)$ is added for each atom~$A$ occurring positively.
Even though such disjunctions are tautological,
they are important under answer set semantics (cf.\ Section~\ref{subsec:semantics}),
as they permit deriving an arbitrary subset of atoms from an aggregate in the head
(which explains the name  ``choice semantics'').
Equation \ref{eq:aggr_lu} splits up an aggregate with lower and an upper into the conjunction of two
conjunctively connected parts.
Note that this split agrees with the semtaics in \cite{ferraris05a} independend of the aggregate $\mathrm{op}$ used.
Finally, Equation (\ref{eq:aggr}) reflects that an unconstrained aggregate atom always holds.

\paragraph{Sum Aggregates}
First, we show how monotone aggregates can be translated into a nested expression.
Using this translation we show how to transfom antimonotone sum aggregates and 
also give a translation that treats negative weights.
\footnote{%
  The translation~$\phi$ we provide for \const{sum} aggregates
  assumes that there are no negative weights for literals,
  while the more sophisticated translation in~\cite{ferraris05a} also
  captures negative weights.
  All current solvers work with positive weights only, and thus
  \gringo\ and \lparse\ use a compilation technique~\cite{siniso02a}
  to eliminate negative weights, which are not permitted in
  \lparse's output format~\cite{lparseManual}.
  As this technique may lead to counterintuitive results~\cite{ferraris05a},
  we strongly recommend not to use negative weights within \const{sum} aggregates.}
\begin{align}
  % l sum+ %
  \shiftleft
  \phi[ l~ \const{\#sum}^{+} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]} ] 
  \label{eq:sum_plus}\\
  &= \bigvee\nolimits_{\{i_1,\dots,i_j\}\subseteq\{1,\dots,n\},l\leq w_{i_1}+\dots+w_{i_j}}
   (\phi[L_{i_1}]\wedge\dots\wedge \phi[L_{i_j}]) \notag 
\end{align}
Equation (\ref{eq:sum_plus}) translates a monotone aggregate~$\const{\#sum}^{+}$
with a lower bound and only positive weights into a nested expressions.

\begin{align}
  % l sum %
  \shiftleft
  \phi[l~\const{\#sum}~\code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]} ]
  \label{eq:sum_neg} \\
  &= \phi [ l + \phi^{+}(w_1) + \dots + \phi^{+}(w_n) ~\const{\#sum}^{+}
    \code{[} \phi^{+}(L_1\code{=}w_1)\code{,} \dots\code{,} \phi^{+}(L_n\code{=}w_n) \code{]} ] \notag \\
  % weight lit %
  \shiftleft
  \phi^{+}(L=w) = 
    \left\{\begin{aligned}
      L=w & \mbox{ if } w \geq 0 \\
      \overline{L}=-w &  \mbox{ otherwise}\\
    \end{aligned}\right. \notag \\
  % weight %
  \shiftleft
  \phi^{+}(w) =-\min\{0,w\} \notag
\end{align}
Equation (\ref{eq:sum_neg}) removes all negative weights from a sum aggregate with a lower bound 
and then applies the transformation described above.
Whenever there is a negative weight, the sign of both literal and weight is inverted and the weight itself is subtracted from the bound.
Note that this transformation produces counter-intuitive results as soon as the aggregate is used recursively.
Nevertheless, there are scenarios where negative weights might be helpfull.
As long as the aggregate is stratified, the transformation corresponds with the intuitive meaning, e.g.,
you can use negative weights to compare two sets $A=\{a,b,c\}$ and $B=\{d,e\}$:
\begin{lstlisting}[numbers=none]
subseteq(A,B) :- #sum [ a=1,b=1,c=1,d=-1,e=-1 ] ] 0.
\end{lstlisting}

\begin{align}
  % sum u %
  \phi[ \const{\#sum} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]}~u ] 
  &= \neg \phi[  u+1~\const{\#sum} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]}]
  \label{eq:sum_u}
\end{align}
Finally, Equation (\ref{eq:sum_u}) translates sum aggregates with an upper bound using the
translation for sum aggregates with lower bounds described above.
Note that again negative literals might cause problems.
Even in the case that all weights are negative (and as such the sum aggregate itself is monotone),
the above translation is used, which might lead to counter-intuitive results.

\paragraph{Max and Min Aggregates} 
We define the sematics of max aggregates using the transformation for sum aggregates described above.
First, we translate max aggregates with upper and lower bound, respectively, into corresponding sum aggregates.
Then we give the semantics of min aggregates using the translation for max aggregates.
Note that in constrast to sum aggregates, max aggregates allow for using non-integer weights as long as there
is some strict total ordering defined on them.
\begin{align}
  % l max
  \shiftleft
  \phi[l~\const{\#max} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]} ]
  \label{eq:max_l}\\
  &= \phi[1~\const{\#sum} \code{[} L_1\code{=}\mathrm{ge}(w_1, l)\code{,} \dots\code{,} L_n\code{=}\mathrm{ge}(w_1, l) \code{]} ]  \notag \\
  % max(w,l)
  \shiftleft
  \mathrm{ge}(a, b) = \left\{\!\begin{aligned}
       1 &\mbox{ if } a \geq b \\
       0 &\mbox{ otherwise}
     \end{aligned}\right. \notag
\end{align}
A (monotone) max aggregate with a lower bound is translated into a sum by setting all weights 
that are greater equal to the lower bound to one and requiring that the sum of weights is at least one.
This transformation preserves the monotonicity of the aggregate and holds independendly of the type of weights used.

\begin{align}
  % max u %
  \shiftleft
  \phi[\const{\#max} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]} u]
  \label{eq:max_u} \\
  &= \phi[\const{\#sum} \code{[} L_1\code{=}\mathrm{ge}(w_1, l)\code{,} \dots\code{,} L_n\code{=}\mathrm{ge}(w_1, l) \code{]} 0]  \notag \\
  \shiftleft
  \mathrm{gt}(a, b) = \left\{\!\begin{aligned}
       1 &\mbox{ if } a > b \\
       0 &\mbox{ otherwise}
     \end{aligned}\right. \notag
\end{align}
Similarly to Equation~(\ref{eq:max_l}), an (antimonotone) max aggregate with an upper bound can be translated into a sum with an upper bound.
Again the monotonicity of the aggregate is preserved.

\begin{align}
  % min u %
  \shiftleft
  \phi[ \const{\#min} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]}~u ] 
  \label{eq:min_u}  \\
  &= \phi[1~\const{\#sum} \code{[} L_1\code{=}\mathrm{ge}(l,w_1)\code{,} \dots\code{,} L_n\code{=}\mathrm{ge}(l,w_1) \code{]}]
  \notag \\
  % l min %
  \shiftleft
  \phi[ l~ \const{\#min} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]} ]
  \label{eq:min_l} \\
  &= \phi[\const{\#sum} \code{[} L_1\code{=}\mathrm{gt}(l,w_1)\code{,} \dots\code{,} L_n\code{=}\mathrm{gt}(l,w_1) \code{]} 0]
  \notag
\end{align}
Analogous to the definition of max aggregates, 
Equations (\ref{eq:min_l}) and (\ref{eq:min_u}) define the semantics of min aggregates.

\paragraph{Average Aggregates}
Avergate can be directly translated into sum aggregates by modifying the weights of its literals.
To see how the translation works consider:
\begin{align*}
\mathrm{avg}\code{[}a_1,\dots,a_n\code{]} &\prec b \\
a_1+\dots+a_n               &\prec n \cdot b \\
a_1 - b + \dots + a_n - b   &\prec 0
\end{align*}
where $\prec$ is some relation.
To avoid problems with empty average aggregates we define that $\mathrm{avg}\code{[} \code{]} \prec b$ iff $0 \prec 0$.
Thus the translation can turn an average aggregate into a sum by subtracting the bound from all its weights.
A strong point of this translation is that it preserves the monotonicity of the average aggregate.
This translation is applied in Equations (\ref{eq:avg_l}) and (\ref{eq:avg_u}).

\begin{align}
  % l avg %
  \shiftleft
  \phi[ l~\const{\#avg} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]} ] 
  \label{eq:avg_l} \\
  &= \phi[ \const{\#sum} \code{[} L_1\code{=}w_1-l \code{,} \dots\code{,} L_n\code{=}w_n-l \code{]}~0 ] 
  \notag \\
  % avg u %
  \shiftleft
  \phi[ \const{\#avg} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]}~u ] = 
  \label{eq:avg_u} \\
  &= \phi[ \const{\#sum} \code{[} L_1\code{=}w_1-u \code{,} \dots\code{,} L_n\code{=}w_n-u \code{]}~0 ]
  \notag
\end{align}

\paragraph{Parity Aggregates}
We also support the two parity aggregates $\code{\#even}$ and $\code{\#odd}$.
Intuitively, an even or odd aggregate is true if the sum of its weights modulo two is one or zero, respectively.
Note that we do not write bounds
Even and odd aggregates are transformed into nested expressions:
\begin{align}
  % odd %
  \shiftleft
  \phi[ \const{\#odd} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]} ] \\
  &=\left\{\begin{aligned}
    \bot & \mbox{ if } n = 0 \\
    \phi[ \const{\#odd} \code{[} L_2\code{=}w_2-u \code{,} \dots\code{,} L_n\code{=}w_n-u \code{]}~0 ] & \mbox{ if } 2 | w_1  \\
    \neg L_1 \wedge \phi[ \const{\#odd} \code{[} L_2\code{=}w_2 \code{,} \dots\code{,} L_n\code{=}w_n \code{]}~0 ] & \\
    \vee L_1 \wedge \neg \phi[ \const{\#odd} \code{[} L_2\code{=}w_2 \code{,} \dots\code{,} L_n\code{=}w_n \code{]}~0 ] & \mbox{ otherwise}
  \end{aligned}\right. \notag
\end{align}

\begin{align}
  % even %
  \shiftleft
  \phi[ \const{\#even} \code{[} L_1\code{=}w_1\code{,} \dots\code{,} L_n\code{=}w_n \code{]} ] \\
  &=\left\{\begin{aligned}
    \top & \mbox{ if } n = 0 \\
    \phi[ \const{\#even} \code{[} L_2\code{=}w_2-u \code{,} \dots\code{,} L_n\code{=}w_n-u \code{]}~0 ] & \mbox{ if } 2 | w_1  \\
    \neg L_1 \wedge \phi[ \const{\#even} \code{[} L_2\code{=}w_2 \code{,} \dots\code{,} L_n\code{=}w_n \code{]}~0 ] & \\
    \vee L_1 \wedge \neg \phi[ \const{\#even} \code{[} L_2\code{=}w_2 \code{,} \dots\code{,} L_n\code{=}w_n \code{]}~0 ] & \mbox{ otherwise}
  \end{aligned}\right. \notag
\end{align}

Item~3--8 specify these parts individually for~\const{sum},~\const{max}, and~\const{min}
aggregates, respectively.
Provided that only non-negative weights are used within \const{sum}s,
all three aggregates are subject to one comparison that is monotone,
meaning that by default it evaluates to \textit{false} so that truth must be established.
For \const{sum} and \const{max} aggregates, these monotone comparisons are related
to lower bounds (Item~3 and~5),
while an upper bound behaves monotonically for \const{min} (Item~8).
In turn, the opposite comparison is antimonotone, that is,
by default it evaluates to \textit{true} unless too many or improper literals hold.
This is the case for comparisons to upper bounds with aggregates
\const{sum} and \const{max} (Item~4 and~6), while it concerns 
the lower bound for \const{min} (Item~7).
% <cut here>
Beyond this special treatment in rule heads,
the truth values of aggregate atoms result from their literals as
may be intuitively expected,
and Item~1--8 basically do nothing but formalizing the standard meanings
of aggregate operations.
Note that the above translations merely define the semantics of aggregates,
which does not imply that solvers unfold them in this way.
In fact, \clasp\ natively supports aggregates without translating them.

As regards syntactic representation,
weight~\const{1} is considered a default,
so that $L_i\code{=}\const{1}$ can simply be written as $L_i$.
For instance, the following (multi)sets of (weighted) literals are the same
when combined with any kind of aggregate operation and bounds:
\begin{lstlisting}[numbers=none,escapechar=@]
[a=1, not b=1, c=2]  @\textnormal{and}@
[a,   not b,   c=2]@\textnormal{.}@
\end{lstlisting}
%
Furthermore,
keyword \const{sum} may be omitted, which in a sense makes \const{sum} the default
aggregate operation.
In fact, the following aggregate atoms are synonyms:
\begin{lstlisting}[numbers=none,escapechar=@]
2 sum [a, not b, c=2] 3  @\textnormal{and}@
2     [a, not b, c=2] 3@\textnormal{.}@
\end{lstlisting}
By omitting keyword \const{sum}, we obtain the same notation as the one of 
so-called ``weight constraints''~\cite{siniso02a,lparseManual}, which are actually aggregate atoms
whose operation is addition.

It is important to note that the (weighted) literals within an aggregate
belong to a multiset.
In particular, if there are multiple occurrences
$L\code{=}w_1,\dots,L\code{=}w_k$
of a literal~$L$, in combination with \const{min} and \const{max},
it is not the same like having $L\code{=}w_1+\dots+w_k$.
To see this, note that the program consisting of the facts:
\begin{lstlisting}[numbers=none]
2 max [a=2].  2 min [a=2].
\end{lstlisting}
has $\{\code{a}\}$ as its unique answer set, while there is no answer set for:
\begin{lstlisting}[numbers=none]
2 max [a,a].  2 min [a,a].
\end{lstlisting}

If literals ought not to be repeated,
we can use \const{count} instead of \const{sum}.
Syntactically, \const{count} requires curly instead of square
brackets, and there must not be any weights within a \const{count} aggregate.
Regarding semantics,
% \begin{equation*}%\label{eq:aggregate:atom}
$(
  l~ 
  \const{count} 
  \code{\char`\{}
  L_1\code{,}
  \dots\code{,}
  L_n
  \code{\char`\}}
  u
)$ 
%\end{equation*}
reduces to
$(
  l~ 
  \const{sum} 
  \code{[}
  L_1\code{=}\const{1}\code{,}
  \dots\code{,}
  L_m\code{=}\const{1}
  \code{]}
  u
)$,
where $\{L_1,\dots,L_m\}=\{L_i \mid 1\leq i\leq n\}$ is obtained by dropping
repeated literals.
Of course, the use of~$l$ and~$u$ is optional also with \const{count}.
As an example, note that the next aggregate atoms express the same:
\begin{lstlisting}[numbers=none,escapechar=@]
1 sum   [a=1, not b=1]     1  @\textnormal{and}@ 
1 count {a,a, not b,not b} 1@\textnormal{.}@  
\end{lstlisting}
Keyword \const{count} can be omitted (like \const{sum}),
so that the following are synonyms:
\begin{lstlisting}[numbers=none,escapechar=@]
1 count {a, not b} 1  @\textnormal{and}@ 
1       {a, not b} 1@\textnormal{.}@  
\end{lstlisting}
The last notation is similar to the one of so-called
``cardinality constraints''~\cite{siniso02a,lparseManual},
which are aggregate atoms using counting as their operation.

After considering the syntax and semantics of ground aggregate atoms,
we now turn our attention to non-ground aggregates.
Regarding contained variables, an atom occurring in an aggregate behaves
similar to an atom on the left-hand side of a condition
(cf.\ Section~\ref{subsec:gringo:condition}).
That is, any variable occurring within an aggregate is a priori local,
and it must be bound via a variable of the same name that is global
or that occurs on the right-hand side of a condition
(with the atom containing the variable in front).
As with local variables of conditions, global variables take priority
during grounding, so that the names of local variables must be chosen
with care to avoid accidental clashes.
Beyond conditions (which are more or less the natural construct to use
for instantiating variables within an aggregate),
classical negation (cf.\ Section~\ref{subsec:gringo:negation}),
built-in arithmetic functions (cf.\ Section~\ref{subsec:gringo:arith}),
intervals (cf.\ Section~\ref{subsec:gringo:interval}), and 
pooling (cf.\ Section~\ref{subsec:gringo:pool})
can be incorporated as usual within aggregates,
where intervals and pooling are expanded locally.%
\footnote{%
  Assignments (cf.\ Section~\ref{subsec:gringo:assign}) and (currently)
  also built-in comparison predicates (cf.\ Section~\ref{subsec:gringo:comp})
  are permitted on the right-hand sides of conditions only.}
That is, an interval gives rise to multiple literals connected via~``\code{,}''
within the same aggregate.
The same applies to pooling in front of a condition,
while it turns into a composite condition chained by~``\code{:}'' on the right-hand side.
The notions of level-restrictedness and stratification
(cf.\ Section~\ref{subsec:lambda} and~\ref{subsec:strat}) are extended
to programs with aggregates by
considering the predicates of all atoms (except for those on the right-hand side of a condition)
in an aggregate atom occurring as the head of a rule~$r$,%
\footnote{%
  The rules in a (maximal) stratified subprogram (cf.\ Section~\ref{subsec:strat})
  cannot have aggregate atoms in the head.}
and by assuming $A\in\nbody{r}$ for all atoms~$A$ of an aggregate in the body of~$r$.
Finally, note that aggregates without bounds are also permitted on the
right-hand sides of assignments, but using this feature is only recommended
for aggregates whose atoms belong to domain predicates because space blow-up
can become a bottleneck otherwise.
The following example, making exhaustive use of aggregates,
nonetheless demonstrates this and other features.

\begin{example}\label{ex:aggr}
Consider a situation where an informatics student
wants to enroll for a number of courses at the beginning of a new term.
In the university calendar, eight courses are found eligible,
and they are represented by the following facts:
%
\lstinputlisting[xrightmargin=-20pt,lastline=8]{examples/aggr.lp}
%
In an instance of \const{course}/$3$,
the first argument is a number identifying one of the eight courses,
and the third argument provides the course's contact hours per week.
The second argument stands for a subject area:
\const{1} corresponding to ``theoretical informatics,''
\const{2}               to ``practical informatics,''
\const{3}               to ``technical informatics,''
and~\const{4}               to ``applied informatics.''
For instance, atom \code{\const{course}(\const{1},\const{2},\const{5})}
expresses that course~\const{1} accounts for~\const{5} contact hours per week
that may be credited to subject area~\const{2} (``practical informatics'').
Observe that a single course is usually eligible for multiple
subject areas.

After specifying the above facts,
the student starts to provide personal constraints on the courses to enroll.
The first condition is that~\const{3} to~\const{6} courses should be enrolled:
%
\lstinputlisting[nolol,firstline=11,lastline=11,firstnumber=9]{examples/aggr.lp}
%
Instantiating the above \const{count} aggregate
yields the following ground rule:%
\marginlabel{%
  The full ground program is obtained by invoking:\\
  \code{\mbox{~}gringo -t \textbackslash\\\mbox{~}examples/aggr.lp}}
%
\begin{lstlisting}[firstnumber=9,stepnumber=9]
3 { enroll(1), enroll(2), enroll(3), enroll(4), 
    enroll(5), enroll(6), enroll(7), enroll(8) } 6.
\end{lstlisting}
%
Observe that an instance of atom~\code{\const{enroll}(\var{C})} is included for each
instantiation of~\var{C} such that \code{\const{course}(\var{C},\var{S},\var{H})}
holds for some values of~\var{S} and~\var{H}.
Duplicates resulting from distinct values for~\var{S} are removed, thus,
obtaining the above set of ground atoms.

The next constraints of the student regard the subject areas of enrolled courses:
%
\lstinputlisting[nolol,firstline=13,lastline=15,firstnumber=10]{examples/aggr.lp}
%
Each of the three integrity constraints above contains a \const{sum} aggregate,
using default weight~\const{1} for literals.
Recalling that \const{sum} aggregates operate on multisets,
duplicates are not removed.
Thus, the integrity constraint in Line~10 is instantiated as follows:
%
\begin{lstlisting}[firstnumber=10,stepnumber=10]
:- [ enroll(1) = 1, enroll(1) = 1, 
     enroll(2) = 1, enroll(2) = 1,
     enroll(3) = 1, enroll(3) = 1,
     enroll(4) = 1, enroll(4) = 1, enroll(4) = 1,
     enroll(5) = 1, enroll(5) = 1,
     enroll(6) = 1, enroll(6) = 1,
     enroll(7) = 1, enroll(7) = 1, enroll(7) = 1,
     enroll(8) = 1, enroll(8) = 1 ] 10.
\end{lstlisting}
%
Note that courses~\const{4} and~\const{7} count three times because
they are eligible for three subject areas, viz., there are three
distinct instantiations for~\var{S} in
\code{\const{course}(\var{4},\var{S},\const{3})} and
\code{\const{course}(\var{7},\var{S},\const{4})}, respectively.
Comparing the above ground instance,
the meaning of the integrity constraint in Line~10 is that the 
number of eligible subject areas over all enrolled courses
must be more than~\const{10}.
Similarly, the integrity constraint in Line~11 expresses the requirement
that at most one course of subject area~\const{2} (``practical informatics'')
is not enrolled,
while Line~12 stipulates that the enrolled courses
amount to less than six nominations of
subject area~\const{3} (``technical informatics'') 
or~\const{4} (``applied informatics'').
Also note that, given the facts in Line~1--8,
we could equivalently have used \const{count} rather than \const{sum} in Line~11,
but not in Line~10 and~12.

The remaining constraints of the student deal with contact hours.
To express them, we first introduce an auxiliary rule and a fact:
%
\lstinputlisting[nolol,firstline=17,lastline=18,firstnumber=13]{examples/aggr.lp}
%
The rule in Line~13 projects instances of \pred{course}/$3$ to
\pred{hours}/$2$, thereby, dropping courses' subject areas.
This is used to not consider the same course multiple times within the following
integrity constraints:
%
\lstinputlisting[nolol,firstline=20,lastline=22,firstnumber=15]{examples/aggr.lp}
%
As Line~15 shows,
we may use default negation via ``\code{not}'' in front of aggregate atoms,
and bounds may be specified in terms of variables.
In fact, by instantiating~\var{M} to~\const{20},
we obtain the following ground instance of the integrity constraint in Line~15:
%
\begin{lstlisting}[firstnumber=15,stepnumber=15]
:- not 18 [ enroll(1) = 5, enroll(2) = 4, 
            enroll(3) = 6, enroll(4) = 3,
            enroll(5) = 4, enroll(6) = 2,
            enroll(7) = 4, enroll(8) = 5 ] 20.
\end{lstlisting}
%
The above integrity constraint states that the \const{sum} of contact hours per week
must lie in-between~\const{18} and~\const{20}.
Note that the \const{min} and \const{max} aggregates in Line~16 and~17, respectively,
work on the same (multi)set of weighted literals as in Line~15.
While the integrity constraint in Line~16 stipulates that any course to enroll
must include more than~\const{2} contact hours,
the one in Line~17 prohibits enrolling for courses of~\const{6} or more contact hours.
Of course, the last two requirements could also be formulated as follows:
%
\begin{lstlisting}[firstnumber=16]
:- enroll(C), hours(C,H), H <= 2.
:- enroll(C), hours(C,H), H >= 6.
\end{lstlisting}

Finally, the following rules illustrate the use of aggregates within assignments:
%
\lstinputlisting[nolol,firstline=24,lastline=25,firstnumber=18]{examples/aggr.lp}
%
Note that the above aggregates have already been used in Line~9 and~15, respectively,
where keywords \const{count} and \const{sum} have been omitted for convenience.
These keywords can be dropped here too, and we merely include them to show the
more verbose notations of \const{count} and \const{sum} aggregates.
However, the usage of aggregates in the last two lines is different from before,
as they now serve to assign an integer to a variable~\var{N}.
In this context, bounds are not permitted, and so none are provided in Line~18 and~19.
The effect of these two lines is that the student can read off the number of
courses to enroll and the amount of contact hours per week from instances of
\pred{courses}/$1$ and \pred{hours}/$1$ belonging to an answer set.%
\marginlabel{%
  To compute the unique answer set of the program, invoke:\\
  \code{\mbox{~}gringo \textbackslash\\
        \mbox{~}examples/aggr.lp | \textbackslash\\
        \mbox{~}clasp -n 0}\\
  or alternatively:\\
  \code{\mbox{~}clingo -n 0 \textbackslash\\
        \mbox{~}examples/aggr.lp}}
%
In fact, running \clasp\ shows the student that a unique
collection of~\const{5} courses to enroll satisfies all requirements:
the courses~\const{1}, \const{2}, \const{4}, \const{5}, and~\const{7},
amounting to~\const{20} contact hours per week.

Although the above program does not reflect this possibility,
it should be noted that (as has been mentioned in Section~\ref{subsec:gringo:condition})
multiple literals may be connected via ``\code{:}'' in order to construct
composite conditions within an aggregate.
As before, the predicates of atoms on the right-hand side of such conditions
must be either domain predicates or built-in.
Furthermore, the usage of non-domain predicates within an aggregate
on the right-hand side of an assignment (like \pred{enroll}/$1$ in Line~18 and~19 above)
is not recommended in general because the space blow-up may be significant.
\eexample
\end{example}

\begin{newstuff}
	\begin{itemize}
		\item there is also $\#$avg, $\#$times, $\#$even, $\#$odd
		\item bingo will support everything but times
		\item heads of stratified aggregates can bind their variables
		\item a \# has to be put in front of every aggregate $\Rightarrow$ adjust examples
	\end{itemize}
\end{newstuff}