### 2010-10-01 doc/guide/options.tex

\section{Command Line Options}\label{sec:options}

In this section, we briefly describe the meanings of
command line options supported by
\gringo\ (Section~\ref{subsec:opt:gringo}),
\clingo\ (Section~\ref{subsec:opt:clingo}),
\iclingo\ (Section~\ref{subsec:opt:iclingo}), and
\clasp\ (Section~\ref{subsec:opt:clasp}).
Each of these tools display their available options
when invoked with flag \code{--help} or \code{-h}.%
\footnote{%
  Note that our description of command line options
  is based on Version~3.0.x of \gringo, \clingo, and \iclingo\
  as well as Version~1.3.x of \clasp.
  While it is rather unlikely that command line options will
  disappear in future versions,
  additional ones might be introduced.
  We will try to keep this document up-to-date,
  but checking the help information shipped 
  with a new version is always a good idea.}
The approach of distinguishing long options, starting with ``\code{--},''
and short ones of the form ``\code{-\textit{l}},''
where \code{\textit{l}} is a letter,
follows the GNU Coding Standards~\cite{GNUcoding}.
For obvious reasons,
short forms are made available only for the most common (long) options.
Some options, also called flags, do not take any argument,
while others require arguments.
An argument \code{\textit{arg}} is provided to a (long) option \code{\textit{opt}}
by writing 
``\code{--\textit{opt}=\textit{arg}}'' or
``\code{--\textit{opt} \textit{arg}},''
while only
``\code{-\textit{l} \textit{arg}}''
is accepted for a short option \code{\textit{l}}.
For each command line option,
we below indicate whether it requires an argument,
and if so, we also describe its meaning.

\subsection{\gringo\ Options}\label{subsec:opt:gringo}

An abstract invocation of \gringo\ looks as follows:
%
\begin{lstlisting}[numbers=none]
gringo [ options | filenames ]
\end{lstlisting}
%
Note that options and filenames do not need to be passed to \gringo\
in any particular order.
If neither a filename nor an option that makes \gringo\ exit (see below)
is provided, \gringo\ reads from the standard input.
In the following, we list and describe the options accepted by \gringo\
along with their particular arguments (if required):
%
\begin{description}
\item[\code{--help,-h}]~\\
Print help information and exit.
\item[\code{--version,-v}]~\\
Print version information and exit.
%\item[\code{--syntax}]~\\
%Print syntax information (about the input language) and exit.
%\item[\code{--stats}]~\\
%Print (extended) statistic information before termination.
\item[\code{--verbose[=n],-V}]~\\
Print additional (progress) information during computation.
Verbosity level one and two are currently not used by \gringo.
Level three prints internal representations of rules during grounding.
(This may be used to identify either semantic errors in an input program or
 performance bottlenecks.)
\item[\code{--const,-c \textit{c}=\textit{t}}]~\\
Replace occurrences (in the input program) of a constant~\code{\textit{c}}
with a term~\code{\textit{t}}.
\item[\code{--text,-t}]~\\
Output ground program in (human-readable) text format.
\item[\code{--reify}]~\\
Output ground program in form of facts.
(These facts can then be used to, i.e., write a \attach{examples/meta.lp}{meta-interpreter}.)
%\comment{add a section about a meta-interpreter}
\item[\code{--lparse,-l}]~\\
Output ground program in \lparse's numerical format~\cite{lparseManual}.
\item[\code{--ground,-g}]~\\
Enable lightweight mode for processing a ground input program.
(This option is recommended to 
 omit unnecessary overhead if the input program is already ground,
 but it leads to a syntax error (cf.\ Section~\ref{subsec:error}) otherwise.)
\item[\code{--ifixed=\textit{n}}]~\\
Use \code{\textit{n}} as fix step number if the input program
contains \const{\#cumulative} or \const{\#volatile} statements.
(This option permits the handling of programs written for \iclingo\
 in a traditional single pass computation.)
\item[\code{--ibase}]~\\
Process only the static part (that can be initiated by a \const{\#base} statement)
of an input program.
(This option may be used to investigate the basic setting of a problem
 including some mutable bound.)
\item[\code{--dep-graph=filename}]~\\
This option can be used to get the dependency graph in from
of a dot\footnote{\url{http://www.graphviz.org/}} file of the program.
\item[\code{--shift}]~\\
Removes disjunction by shifting (see \cite{gellif91a}).
\end{description}
%
The default command line when invoking \gringo\ is as follows:
%
\begin{lstlisting}[numbers=none]
gringo --lparse
\end{lstlisting}
That is, \gringo\ usually outputs a ground program in \lparse's numerical format,
dealt with by various ASP solvers \cite{gekanesc07b,siniso02a,linzha04a}.

\subsection{\clingo\ Options}\label{subsec:opt:clingo}

ASP system \clingo\ combines grounder \gringo\ and solver \clasp\
via an internal interface.
An abstract invocation of \clingo\ looks as follows:
%
\begin{lstlisting}[numbers=none]
clingo [ number | options | filenames ]
\end{lstlisting}
%
A numerical argument is permitted for backward compatibility to
the usage of solver \smodels~\cite{siniso02a},
where it specifies the maximum number of answer sets to be computed
(\code{0}~standing for all answer sets).
As with \gringo, a number, options, and filenames do not need to be passed to \clingo\
in any particular order.
Given that \clingo\ combines \gringo\ and \clasp,
it accepts all options described in the previous section
and in Section~\ref{subsec:opt:clasp}.
In particular, (long) options \code{--help} and \code{--version}
make \clingo\ print the desired information and exit,
while \code{--text}, \code{--lparse}, and \code{--reify}
instruct \clingo\ to output a ground program (rather than solving it) like \gringo.
If neither a filename nor an option that makes \clingo\ exit
(see Section~\ref{subsec:opt:gringo}) is provided, \clingo\ reads from the standard input.
Beyond the options described in Section~\ref{subsec:opt:gringo} and~\ref{subsec:opt:clasp},
\clingo\ has a single additional option:
%
\begin{description}
\item[\code{--clasp}]~\\
Run \clingo\ as a plain solver (using embedded \clasp).
\end{description}
%
Finally, the default command line when invoking \clingo\
consists of all \clasp\ defaults (cf.\ Section~\ref{subsec:opt:clasp}).

\subsection{\iclingo\ Options}\label{subsec:opt:iclingo}

Incremental ASP system \iclingo\ extends \clingo\ by interleaving
grounding and solving for problems including a mutable bound.
An abstract invocation of \iclingo\ is as with \clingo:
%
\begin{lstlisting}[numbers=none]
iclingo [ number | options | filenames ]
\end{lstlisting}
%
The external behavior of \iclingo\ is similar to \clingo,
described in the previous section, except for the fact that
option \code{--ifixed} is ignored by \iclingo\ if not run as a grounder
(via one of (long) options \code{--text}, \code{--lparse}, or \code{--reify}).
However, option \code{--clingo} (see below) may be used to let \iclingo\
work like \clingo.
The additional options of \iclingo\ focus on customizing
incremental computations:
%
\begin{description}
\item[\code{--istats}]~\\
Print statistic information for each incremental solving step.
\item[\code{--imin=\textit{n}}]~\\
Perform at least \code{\textit{n}} incremental solving steps before termination.
(This may be used to force steps regardless of the termination condition
 set via \code{--istop}.)
\item[\code{--imax=\textit{n}}]~\\
Perform at most \code{\textit{n}} incremental solving steps before termination.
(This may be used to limit steps regardless of the termination condition
 set via \code{--istop}.)
\item[\code{--istop=SAT|UNSAT}]~\\
Terminate after an incremental solving step 
in which some (\code{SAT}) or no (\code{UNSAT})
answer set has been found.
\item[\code{--iquery=\textit{n}}]~\\
Start with incremental solving at step number~\code{\textit{n}}.
(This may be used to skip some solving steps,
 still accumulating static and cumulative rules for these steps.)
\item[\code{--ilearnt=keep|forget}]~\\
Maintain (\code{keep}) or delete (\code{forget}) learnt constraints
in-between incremental solving steps.
(This option configures the behavior of embedded \clasp.)
\item[\code{--iheuristic=keep|forget}]~\\
Maintain (\code{keep}) or delete (\code{forget}) heuristic information
in-between incremental solving steps.
(This option configures the behavior of embedded \clasp.)
\item[\code{--clingo}]~\\
Run \iclingo\ as a non-incremental ASP system (like \clingo).
\end{description}
%
As with \clingo, the default command line when invoking \iclingo\
consists of all \clasp\ defaults, explained in the next section,
along with \code{--istop=SAT}, \code{--iquery=1}, \code{--ilearnt=keep}, and
\code{--iheuristic=forget}.
That is, incremental solving starts at step number~\code{1} and stops
after a step in which some answer set has been found.
In-between incremental solving steps, embedded \clasp\ maintains learnt constraints
but deletes heuristic information.

\subsection{\clasp\ Options}\label{subsec:opt:clasp}

Stand-alone \clasp~\cite{gekanesc07b} is a solver for ground programs in
\lparse's numerical format~\cite{lparseManual}. Beyond that, it can also
be used as a SAT (on a simplified version of DIMACS/CNF format%
\footnote{%
\url{http://www.satcompetition.org/2009/format-benchmarks2009.html}
}), or PB solver (on OPB format%
\footnote{%
\url{http://www.cril.univ-artois.fr/PB09/solver_req.html}
}).
An abstract invocation of \clasp\ looks as follows:
%
\begin{lstlisting}[numbers=none]
clasp [ number | options | filename ]
\end{lstlisting}
%
As with \clingo\ and \iclingo,
a numerical argument specifies the maximum number of answer sets to be computed,
where \code{0}~stands for all answer sets.
(The number of requested answer sets can likewise be set via long option
 \code{--number} or its short form~\code{-n}.)
If neither a filename nor an option that makes \clasp\ exit (see below) is provided,
\clasp\ reads from the standard input.%
\footnote{%
 In earlier versions of \clasp\ filenames had to be given via long option \code{--file} 
 or its short form \code{-f}.}
In fact, it is typical to use \clasp\ in a pipe with \gringo\ in the following way:
\begin{lstlisting}[numbers=none]
gringo [ ... ] | clasp [ ... ]
\end{lstlisting}
In such a pipe, \gringo\ instantiates an input program and outputs the ground
rules in \lparse's numerical format,
which is then consumed by \clasp\ that computes and outputs answer sets.
Note that \clasp\ offers plenty of options to configure its behavior.
We thus categorize them according to their functionalities in the following description.

\subsubsection{General Options}\label{subsec:clasp:general}

We below group general options of \clasp,
used to configure its global behavior.
%
\begin{description}
\item[\code{--help,-h}]~\\
Print help information and exit.
\item[\code{--version,-v}]~\\
Print version information and exit.
\item[\code{--pre}]~\\
Run ASP preprocessor then print preprocessed input program and exit.
\item[\code{--verbose[=n],-V}]~\\
Configure printing of (progress) information during computation. 
Argument $\code{\textit{n}}=\code{0}$ disables progress information, while
$\code{\textit{n}}=\code{1}$ prints basic, and $\code{\textit{n}}=\code{2}$
extended information.
\item[\code{--stats}]~\\
Print (extended) statistic information before termination.
\item[\code{--number,-n \textit{n}}]~\\
Compute at most~\code{\textit{n}} answer sets,
$\code{\textit{n}}=\code{0}$ standing for compute all answer sets.
\item[\code{--quiet,-q}]~\\
Do not print computed answer sets.
(This is useful for benchmarking.)
\item[\code{--time-limit=\textit{t}}]~\\
Force termination after \code{\textit{t}} seconds.
\item[\code{--search-limit=\textit{n},\textit{m}}]~\\
Force termination after either \code{\textit{n}} 
conflicts or \code{\textit{m}} restarts.
\item[\code{--seed=\textit{n}}]~\\
Use seed~\code{\textit{n}} (rather than~\code{1}) for random number generator.
\item[\code{--solution-recording}]~\\
Switch from backtrack-based enumeration~\cite{gekanesc07c} to
enumeration based on solution recording. 
Note that this  mode is prone to blow up 
in space in view of an exponential number of solutions in the worst case.
\item[\code{--restart-on-model}]~\\
Restart the search from scratch (rather than doing enumeration~\cite{gekanesc07c})
after finding an answer set. Note: This mode implies solution recording and hence
has the same caveat. It is mainly useful when searching for an optimal 
solution because, first, restarting may greatly speed up finding the optimum, and second,
only the current optimum needs to be recorded.
\item[\code{--project}]~\\
Project answer sets to named atoms and only enumerate unique projected solutions~\cite{gekasc09a}.
\item[\code{--brave}]~\\
Compute the brave consequences (union of all answer sets) of a logic program.
\item[\code{--cautious}]~\\
Compute the cautious consequences (intersection of all answer sets) of a logic program.
\item[\code{--opt-all}]~\\
Compute all optimal answer sets (cf.\ Section~\ref{subsec:gringo:optimize}).
(This is implemented by enumerating~\cite{gekanesc07c} answer sets that are not worse
 than the best one found so far.)
\item[\code{--opt-value=\textit{n1}[,\textit{n2},\textit{n3}...]}]~\\
Initialize objective function(s) to minimize with \code{\textit{n1}[,\textit{n2},\textit{n3}...]}.
\item[\code{--opt-ignore}]~\\
Ignore any optimize statements of a logic program during computation.
\item[\code{--opt-heu}]~\\
Consider optimize statements in heuristics.
\item[\code{--supp-models}]~\\
Compute supported models~\cite{apblwa88a} (rather than answer sets).
\item[\code{--trans-ext=all|choice|weight|dynamic|no}]~\\
Compile extended rules~\cite{siniso02a} into normal rules of form (\ref{eq:normal:rule}).
Arguments \code{choice} and \code{weight}
state that all ``choice rules'' or all ``weight rules,'' respectively,
are to be compiled into normal rules,
while \code{all} means that both and \code{no} that none of them
are subject to compilation. If argument \code{dynamic} is given, 
\clasp\ heuristically decides whether or not to compile individual
``weight rules''.
\item[\code{--eq=\textit{n}}]~\\
Run equivalence reasoning~\cite{gekanesc08a} for \code{\textit{n}} iterations,
$\code{\textit{n}}=\code{-1}$ and $\code{\textit{n}}=\code{0}$ standing for
run to fixpoint or do not run equivalence reasoning, respectively.
\item[\code{--backprop}]~\\
Enable backpropagation in ASP-preprocessing.
\item[\code{--sat-prepro=yes|no|\textit{n1}[,\textit{n2},\textit{n3}]}]~\\
Run \emph{SatElite}-like preprocessing~\cite{eenbie05a}
for at most~\code{\textit{n1}} iterations 
($\code{\textit{n1}}=\code{-1}$ standing for run to fixpoint),
using cutoff~\code{\textit{n2}} for variable elimination
($\code{\textit{n2}}=\code{-1}$ standing for no cutoff), and
for no longer than~\code{\textit{n3}} seconds
($\code{\textit{n3}}=\code{-1}$ standing for no time limit).
Arguments \code{yes} and \code{no} mean
$\code{\textit{n1}}=\code{\textit{n2}}=\code{\textit{n3}}=\code{-1}$
(that is, run to fixpoint)
or that \emph{SatElite}-like preprocessing is not to be run at all, respectively.
\end{description}
%
Having introduced the general options of \clasp,
let us note that the options below \code{--supp-models} in the above list
are quite low-level and more or less an issue of fine-tuning.
More important is the fact that virtually all optimization functionalities
are only provided by \clasp\ if the maximum number of answer sets to be
computed is set to~\code{0} (standing for all answer sets),
as it is likely to stop search too early otherwise.
The same applies to computing either brave or cautious consequences
(via one of the flags \code{--brave} and \code{--cautious}).

\subsubsection{Search Options}\label{subsec:clasp:search}

The options listed below can be used to configure the main search
strategies of \clasp.
%
\begin{description}
\item[\code{--lookahead=atom|body|hybrid|no}]~\\
Apply failed-literal detection~\cite{freeman95}
to atoms (with argument \code{atom}),
to rule bodies (with argument \code{body}), or
to atoms and rule bodies like in \nomorepp~\cite{angelinesc05c} (with argument \code{hybrid}).
Failed-literal detection is switched off via argument~\code{no}.
\item[\code{--initial-lookahead=\textit{n}}]~\\
Apply failed-literal detection in preprocessing,
and for the first \code{\textit{n}} decisions during search.
\item[\code{--heuristic=Berkmin|Vmtf|Vsids|Unit|None}]~\\
Use \emph{BerkMin}-like decision heuristic~\cite{golnov02a}
(with argument~\code{Berkmin}),
\emph{Siege}-like decision heuristic~\cite{ryan04a}
(with argument~\code{Vmtf}),
\emph{Chaff}-like decision heuristic~\cite{momazhzhma01a}
(with argument~\code{Vsids}),
\emph{Smodels}-like decision heuristic~\cite{siniso02a}
(with argument~\code{Unit}),
or (arbitrary) static variable ordering
(with argument~\code{None}).
\item[\code{--rand-freq=\textit{p}}]~\\
Perform random (rather than heuristic) decisions
with probability $0\leq\code{\textit{p}}\leq 1$.
\item[\code{--rand-prob=yes|no|\textit{n1},\textit{n2}}]~\\
Run \emph{Satzoo}-like random probing~\cite{een03a},
initially performing~\code{\textit{n1}} passes of up to~\code{\textit{n2}} conflicts
making random decisions.
Arguments \code{yes} and \code{no} mean
$\code{\textit{n1}}=\code{50},\linebreak[1]\code{\textit{n2}}=\code{20}$
or that random probing is not to be run at all, respectively.
\item[\code{--rand-watches=yes|no}]~\\
Initially choose watched literals randomly (with argument \code{yes})
or systematically (with argument \code{no}).
\end{description}

\subsubsection{Lookback Options}\label{subsec:clasp:back}

The following options have an effect only if lookback techniques
are turned on, that is, option \code{--no-lookback} is not used.
%
\begin{description}
\item[\code{--no-lookback}]~\\
Disable all lookback techniques.
This option is included mainly for comparison purposes, and its use
is not generally recommended.
\item[\code{--restarts,-r \textit{n1}[,\textit{n2},\textit{n3}]|no}]~\\
Choose and parameterize a restart policy.
If a single argument \code{\textit{n1}} is provided,
\clasp\ restarts search from scratch after a number of conflicts
determined by a universal sequence~\cite{lusizu93a}, where \code{\textit{n1}}
constitutes the base unit.
If two arguments \code{\textit{n1},\textit{n2}} are specified,
\clasp\ runs a geometric sequence~\cite{eensor03a},
restarting every $\code{\textit{n1}}*\code{\textit{n2}}^i$ conflicts,
where~$i$ is the number of restarts performed so far.
Given three arguments \code{\textit{n1},\textit{n2},\textit{n3}},
\clasp\ repeats geometric restart sequence $\code{\textit{n1}}*\code{\textit{n2}}^i$
when it reaches an outer limit~$\code{\textit{n1}}*\code{\textit{n3}}^j$~\cite{biere08a},
where~$j$ counts how often the outer limit has been hit so far.
Finally, restarts are disabled via argument~\code{no}.
\item[\code{--local-restarts}]~\\
Count conflicts locally~\cite{ryvstr08a} (rather than globally)
for deciding when to restart.
\item[\code{--bounded-restarts}]~\\
Perform bounded restarts in answer set enumeration~\cite{gekanesc07c}.
\item[\code{--reset-restarts}]~\\
Reset restart strategy whenever an answer set is found.
\item[\code{--save-progress[=\textit{n}]}]~\\
Use cached (rather than heuristic) decisions~\cite{pipdar07a} if available.
Cache decisions on backjumps $>$ \code{\textit{n}}.
\item[\code{--shuffle,-s \textit{n1},\textit{n2}}]~\\
Shuffle internal data structures after~\code{\textit{n1}} restarts
($\code{\textit{n1}}=\code{0}$ standing for no shuffling)
and then reshuffle every~\code{\textit{n2}} restarts 
($\code{\textit{n2}}=\code{0}$ standing for no reshuffling).
\item[\code{--deletion,-d \textit{n1}[,\textit{n2},\textit{n3}]}]~\\
Limit the number of learnt constraints to 
$\min\{(c/\code{\textit{n1}})*\code{\textit{n2}}^i,c*\code{\textit{n3}}\}$,
where~$c$ is the initial number of constraints and~$i$ is the number of restarts performed so far.
\item[\code{--reduce-on-restart}]~\\
Delete a portion of learnt constraints after every restart.
\item[\code{--estimate}]~\\
Base the initial limit of learnt constraints on an estimate
of the problem's complexity.
\item[\code{--strengthen=bin|tern|all|no}]~\\
Check binary (with argument \code{bin}),
binary and ternary (with argument \code{tern}), or 
all (with argument \code{all}) antecedents for
self-subsumption~\cite{eenbie05a}
in order to strengthen a constraint to learn.
Strengthening is disabled via argument~\code{no}.
\item[\code{--recursive-str}]~\\
Recursively apply strengthening, as proposed in~\cite{soreen05a}.
\item[\code{--loops=common|distinct|shared|no}]~\\
Learn loop nogood~\cite{gekanesc07a} per atom in an unfounded set~\cite{gerosc91a}
(with argument \code{common}),
shrink unfounded set before learning another loop nogood
(with argument \code{distinct}),
learn loop formula~\cite{linzha04a} for atoms in an unfounded set
(with argument \code{shared}),
or do not record unfounded sets at all
(with argument \code{no}).
\item[\code{--contraction=\textit{n}}]~\\
Temporarily truncate learnt constraints over more than \code{\textit{n}} variables~\cite{ryan04a}.
\end{description}
%
Let us note that switching the above search and lookback options can have dramatic effects
(both positively and negatively) on the search performance of \clasp.
If performance bottlenecks are observed, it is worthwhile to give 
\code{Vmtf} and \code{Vsids} for \code{--heuristic} a try, in particular,
when the program under consideration is huge 
but scarcely yields conflicts during search.
Furthermore, we suggest trying the universal restart sequence~\cite{lusizu93a}
with different base units~\code{\textit{n1}} or even disabling restarts
(both via (long) option \code{--restarts} or its short form~\code{-r})
in case that performance needs to be improved.
For a brief overview on fine-tuning see \cite{gekasc09b}.
Finally, let us consider the default command line of \clasp:
%
\begin{lstlisting}[numbers=none]
clasp 1 --verbose=1 --seed=1 --trans-ext=no --eq=5
        --sat-prepro=no --lookahead=no
        --heuristic=Berkmin --deletion=3.0,1.1,3.0
        --rand-freq=0.0 --rand-watches=yes --rand-prob=no 
        --restarts=100,1.5 --shuffle=0,0 --strengthen=all
        --loops=common --contraction=250
\end{lstlisting}
%
Considering only the most significant defaults,
numeric argument~\code{1} instructs \clasp\ to terminate immediately
after finding an answer set,
while \code{--restarts=100,1.5} lets \clasp\ apply a geometric
restart policy.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "guide"
%%% End: